### 2021-03-05 ch/4-structure.tex
Incorporate Novi's feedback for chapter 1
Use include instead of import
This has the advantage that I can compile one chapter at a time.

% \chapter{Structure of the thesis}\label{ch:content}
\chapter{Summary of contributions}\label{ch:content}
The main contributions of this thesis are as follows.
All presented approaches rely on some form of side information with allows us to overcome the dataset bias.
This side information is always significantly easier to obtain than unbiased data.

@@

We can interpret the contributions of this chapter in two ways.
The first corresponds to the definition-centric view of dataset bias,
and the second to the groundtruth-centric view.
\begin{enumerate}
  \item
    We can say that the classifier should satisfy demographic parity in its predictions,
    and learning from a balanced training set is just one particular way to achieve this.
    In this view, the pseudo labels have no deeper meaning and are just a computational trick.
  \item
    We can see the training set as a corrupted version of a true dataset, which is balanced (\(y\perp s\)),
    and so, by learning from these pseudo labels, we are simply approximating the true dataset.
    However, we do not actually have access to the true dataset; we only know that it is balanced.
    In order to evaluate the trained model, we compute fairness metrics with respect to \ac{DP}.
\end{enumerate}
Within the chapter, we sometimes jump between these two views.

@@

As an example, consider a dataset where the task is to distinguish smiling from non-smiling faces.
Say, we initially have a very diverse and balanced training set,
but then, from the set of smiling faces, we remove nearly all samples where the person does not have red hair,
and from the set of non-smiling faces, we remove nearly all samples with non-black hair.
The result is that even though the labels are still correct, they do not point reliably to the right target anymore.
As hair colour is easier to recognize with a \ac{CNN}, the labels have effectively been turned into hair colour labels.
In other words, hair colour has become a \emph{spurious attribute}.
In the following, we denote the spurious attribute with \(s\),
as it takes a role that is very similar to that of the sensitive attribute that was also denoted by \(s\).

>>>>

\chapter{Summary of contributions}\label{ch:content}
The main contributions of this thesis are as follows.
All presented approaches rely on some form of side information with allows us to overcome the dataset bias.
This side information is always significantly easier to obtain than unbiased data.

@@

We can interpret the contributions of this chapter in two ways.
The first corresponds to the definition-centric view of dataset bias,
and the second to the ground-truth-centric view.
\begin{enumerate}
  \item
    We can say that the classifier should satisfy demographic parity in its predictions,
    and learning from a balanced training set is just one particular way to achieve this.
    In this view, the pseudo labels have no deeper meaning and are just a computational trick.
  \item
    We can see the training set as a corrupted version of a true dataset, which is balanced (\(y\perp s\)),
    and so, by learning from these pseudo labels, we are simply approximating the true dataset.
    However, we do not actually have access to the true dataset; we only know that it is balanced.
    In order to evaluate the trained model, we compute fairness metrics with respect to \ac{DP}.
\end{enumerate}
Within the chapter, we sometimes jump between these two views.

@@

As an example, consider a dataset where the task is to distinguish smiling from non-smiling faces.
Say, we initially have a very diverse and balanced training set,
but then, from the set of smiling faces, we remove nearly all samples where the person does not have red hair,
and from the set of non-smiling faces, we remove nearly all samples with non-black hair.
The result is that even though the labels are still correct, they do not point reliably to the right target anymore.
As hair colour is easier to recognise with a \ac{CNN}, the labels have effectively been turned into hair colour labels.
In other words, hair colour has become a \emph{spurious attribute}.
In the following, we denote the spurious attribute with \(s\),
as it takes a role that is very similar to that of the sensitive attribute that was also denoted by \(s\).

### 2021-03-05 ch/6-discussion.tex
Use include instead of import
This has the advantage that I can compile one chapter at a time.

Thus, correcting dataset bias remains a challenging topic
and one that is relevant to today's machine learning applications.
Any cutting-edge \ac{ML} system will have to deal with imperfect data,
especially if the collected data is human-related.
The possible effects of these data imperfections are certainly highly undesirable:
a photo tagging service might only work for a certain kind of person;
a speech recognition system might only work for a certain kind of dialect.
If, in these situations, enough representative (but unlabeled) data is available,
then the methods presented here can be used to try and correct the problem.

One area of machine learning that has recently seen increased interest is unsupervised learning.
The latter two chapters also make use of it to some degree.
The exciting promise of unsupervised learning in general is
that labor-intensive labeling is not needed and so vast amounts of existing, unlabeled data can be put to good use.
One could ask the question whether bias-correcting methods are still needed, with access to so much data.
It could be that, while the data is certainly not perfect,
there is so much of it that the biased parts ``cancel each other out''.
However, recent investigations of the GPT models do not seem to bear this out.
One reason for this might be the way these models are trained at the moment:
they maximize the probability assigned to the next token.
Thus, such a model has to account for the wide array of human opinions and assign a non-zero probability to all of them.
So, when asked to summarize a text, GPT does not give the \emph{best} summary;
it will give a summary that an average person might have written.
However, with the help of a very high quality labeled dataset (that was expensive to create),
GPT could be finetuned to actually produce very good summaries.
I suspect this pattern of learning the basics in an unsupervised fashion
and then finetuning with high-quality labels will continue in the future.

>>>>

Thus, correcting dataset bias remains a challenging topic
and one that is relevant to today's machine learning applications.
Any cutting-edge \ac{ML} system will have to deal with imperfect data,
especially if the collected data is human-related.
The possible effects of these data imperfections are certainly highly undesirable:
a photo tagging service might only work for a certain kind of person;
a speech recognition system might only work for a certain kind of dialect.
If, in these situations, enough representative (but unlabelled) data is available,
then the methods presented here can be used to try and correct the problem.

One area of machine learning that has recently seen increased interest is unsupervised learning.
The latter two chapters also make use of it to some degree.
The exciting promise of unsupervised learning in general is
that labor-intensive labeling is not needed and so vast amounts of existing, unlabelled data can be put to good use.
One could ask the question whether bias-correcting methods are still needed, with access to so much data.
It could be that, while the data is certainly not perfect,
there is so much of it that the biased parts ``cancel each other out''.
However, recent investigations of the GPT models do not seem to bear this out.
One reason for this might be the way these models are trained at the moment:
they maximise the probability assigned to the next token.
Thus, such a model has to account for the wide array of human opinions and assign a non-zero probability to all of them.
So, when asked to summarise a text, GPT does not give the \emph{best} summary;
it will give a summary that an average person might have written.
However, with the help of a very high quality labelled dataset (that was expensive to create),
GPT could be finetuned to actually produce very good summaries.
I suspect this pattern of learning the basics in an unsupervised fashion
and then finetuning with high-quality labels will continue in the future.
